
\section{Deliverables}

The proposed project includes the following deliverables: 

\begin{enumerate}

%%%what does this mean???
%{\em Systematic experimental paradigms\/} 
\item An {\em experimental platform\/} in which humans can interact
  with artificial agents, and that thereby provides a space for
  artificial agents to demonstrate their capacity to represent, learn,
  and ultimately activate norms in context-appropriate ways.

\item {\em Software infrastructure for learning together with
  machine-learning algorithms\/} that enable the representation,
  learning, and activation of norms in simulated stochastic games.
  Code will be made available as a part of the Brown-UMBC
  Reinforcement Learning and Planning
  library\footnote{http://burlap.cs.brown.edu}.

\item Data comprised of the behaviors produced by both humans and
  machines in stochastic games, under varying conditions of complexity
  and uncertainty, as well empirical evaluations of those behaviors.
  These data will form a set of {\em computational benchmarks}, which
  other researchers can freely use to evaluate their own computational
  models of norms.

\item All of the above for coordination problems, specifically.  This
  will produce a set of {\em coordination benchmark}, which we hope
  will inspire other researchers to explore alternative approaches to
  human-machine collaboration.

\end{enumerate}

In year two, we will organize a workshop on the machine learning of
social norms. Such a gathering is timely, as a number of subareas are
grappling with related problems. We will include researchers in
human-robot collaboration~\cite{gopalan15}, intelligent software,
end-user programming of household devices~\cite{ur14}, agent-based
bidding~\cite{tac:book}, multi-agent reinforcement
learning~\cite{sodomka13}, computational social
cognition~\cite{baker14}, and others.
