
\section{Broader Impacts}

The many devices in our world that collect and share data among the
ever-growing ``Internet of Things" cannot yet intelligently coordinate
their actions to carry out complex tasks without extensive human
programming and configuration.  As this collection of devices grows
and begins to include more traditional robotic actors and more
sophisticated software agents, the need for these entities to not just
communicate but work collaboratively is only going to increase.  We
will want and need them to work together with and without us to
accomplish everyday tasks and help us make better decisions.

The domains that would be positively impacted by machines
that can collaborate effectively with humans is endless, such as robotic sous chefs and
gerontechnological support for aging in place.
%
In particular, the results of this project synergize with the main efforts of 
Brown University's Humanity Centered Robotics Initiative (HCRI), of
which co-PI Littman is co-director. Specifically, within HCRI there
are ongoing efforts to design robotic systems that support independent
living tasks. For an elderly person to trust and collaborate on tasks with a machine effectively, the machine must act in a manner that the elderly person expects. This project lies the foundation for these important applications.

In an effort to increase diversity in computer science, we have
pre-selected two graduate students to participate in this
project---one woman, and the other Hispanic.  All students (both
graduate and undergraduate) who join our team will benefit from collaborating with the cognitive psychologists on our team.
Specifically, they will strengthen their understanding of a number of
fields, all of which are critical to the development of artificial
agents that collaborate effectively with humans: e.g., behavioral
economics, cognitive psychology, reinforcement learning, and software
engineering.

``One of the things that [President Obama] really strongly believe[s]
in is that we need to have more girls interested in math, science, and
engineering.''  To that end, we also plan to integrate our work on
this project into Artemis, a free summer program run at Brown and
directed by PI Greenwald that introduces rising 9th grade girls to
computational thinking.
%\commentj{Explicit mention that this is a STEM priority?}  
For example, we might have the Artemis girls teach a robot to
collaborate with them on various tasks, such as finding and navigating
to common household objects.
%folding laundry.
%\commenta{insert a different example!!!}
%\commenta{You know, girls should be doing housework.} 
% JLA: yeah, i was concerned about the folding laundry example for that reason... 

Our deliverables include an open-source publicly accessible toolkit for machine norm learning via reinforcement learning. Further, we will build a database of machine-machine, human-machine, and human-human experimental results on norm-learning problems, which can serve as a benchmark for future researchers to build artifical agents that increasingly achieve human-like behavior. 
%
Finally, we expect to publish the results of the proposed research in
top-tier archival, conference proceedings and journals with high
impact factors, and present our work at both computer science and
psychology conferences.




