\documentclass[11pt,letterpaper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{times}

\begin{document}

% \pagestyle{empty}

\begin{center}
{\Large Collaboration Plan: RI:Med:Socially Rational Artificial Agents: \\ \vspace{.5ex} A Key to Human-Machine Collaboration}
\end{center}

\section*{Roles}
% 1) the specific roles of the project participants in all organizations involved;

PIs Greenwald and Littman will work together to direct the
project. Both investigators will jointly manage GRAs supported on the
project and will also share the responsibility for setting research
directions, scheduling milestones, submitting reports, reviewing study
protocols, coordinating IRB submissions/revisions, and targeting
dissemination venues.

\section*{Management}
% 2) information on how the project will be managed across all the investigators, institutions, and/or disciplines;

The project investigators will manage the project using a combination of
regular project meetings, one-on-one student meetings, and periodic
administrative meetings. This structure will enable coordination among the PIs,
and facilitate the joint responsibility for project outcomes. 

\section*{Mechanisms}
% 3) identification of the specific coordination mechanisms that will enable cross-investigator, cross-institution, and/or cross-discipline scientific integration (e.g., yearly workshops, graduate student exchange, project meetings at conferences, use of the grid for videoconferences, software repositories, etc.),

Because all of the project investigators and supported GRAs will be
located in the Computer Science Department at Brown University, the
vast majority of project meetings will take place in person. The PIs
will continue to hold weekly hour-long project meetings in most weeks
with all supported GRAs and project-affiliated undergraduats, as well
as shorter one-on-one meetings with GRAs for professional
development. Additionally, when travel schedules prevent in-person
meetings, video conferencing will substitute. The team will use
freely-available services like Skype and/or Google Hangouts.

To support software development, experimental data collection, and the
protection of sensitive results from human studies, the project team
will use VMs mantained by the department's technical staff.  Files
will be backed up regularly by the technical staff.

PI Greenwald and Littman will additionally hold periodic
administrative meetings to review the project's progress towards
research goals. They will also address drafting of annual reports,
monitoring the burn rate of funds, and broader impacts activities in
those meetings.

\section*{Budgetary Support}
% and 4) specific references to the budget line items that support collaboration and coordination mechanisms.

Because of the single-institution composition of the project
personnel, budgetary support for collaboration is minimal. Funds for
the VM maintenance and university data warehousing have been
allocated. 

% Further, because laptops will be purchased using funds from
% this project to collect data from participants in different locations,
% the built-in cameras will be used for video conferencing when
% in-person meetings aren't possible.

\end{document}
