
\vspace{\up}
\paragraph{Deliverables}

The proposed project includes the following deliverables: 

\begin{enumerate}

\item An {\em experimental platform\/} in which humans can interact
  with artificial agents, thereby providing a space for artificial
  agents in general, and our socially rational agents, specifically to
  demonstrate their capacity to represent, learn, and apply
  collaborative behaviors in context-appropriate ways.

\item {\em Software infrastructure for machine learning algorithm
  development and experimentation\/} that enable the representation
  and learning of collaborative behaviors in simulated stochastic games.
  Code will be made available as a part of the Brown-UMBC
  Reinforcement Learning and Planning library.%
\footnote{The open source Brown-UMBC Reinforcement Learning and
  Planning (BURLAP) library (\url{http://burlap.cs.brown.edu}) is the
  software infrastructure that will be used for algorithm development
  in this project.  Developed by James MacGlashan under the direction
  of co-PI Littman, the system provides a flexible and powerful
  implementation of a wide variety of reinforcement-learning
  algorithms and environments, and has been extensively used in both
  published research and education (including in a
  reinforcement-learning MOOC).  The algorithm which formed the basis
  of the work we describe in this proposal (RHC-MLIRL) is available as
  part of the BURLAP distribution.}

\item Data comprised of the behaviors produced by both humans and
  machines on a test bed of stochastic games, varying the degrees of
  complexity and uncertainty, as well evaluations of those behaviors.
  It is our intent that these data will form a set of {\em
    computational benchmarks}, which other researchers will be free to
  use to evaluate their own approaches to collaborative learning.

\item Syllabus for an undergraduate class on ``Social autonomous
  driving'' in which students learn to endow autonomous vehicles with
  the ability to represent, learn, and apply collaborative driving
  algorithms.

\end{enumerate}

In year two, we plan to organize a workshop (e.g., a AAAI symposium)
on the machine learning of socially rational behavior. Such a
gathering would be timely, as a number of subareas are grappling with
related problems. We will include researchers in human--robot
collaboration~\cite{gopalan15}, 
% intelligent software\commenta{missing reference}, 
self-driving cars~\cite{dolgov2010path}
end-user
programming of household devices~\cite{ur14}, agent-based
bidding~\cite{tac:book}, multi-agent reinforcement
learning~\cite{sodomka13}, computational social
cognition~\cite{baker14}, and others.
% \commenta{do you want to reference self-driving cars here?}
