
\section{Broader Impacts}

Recent trends in computer science and artificial intelligence are
moving us toward increasing dependence on human--computer
collaborative systems in which people and software make decisions that
impact one another.  Some fields, such as human--computer interaction
(HCI), focus on systems in which a human being is primarily in
control, and the computer's decisions are assessed in terms of the
positive impact they have on the user~\cite{maes1993learning}.  An
interesting recent example is Centaur
chess\footnote{\url{http://www.huffingtonpost.com/mike-cassidy/centaur-chess-shows-power_b_6383606.html}}
in which a human and machine team up to play on the same side in
chess, both making decisions, but with the machine acting as an
advisor and the human deciding which moves to actually make.  Other
fields, like crowd sourcing and human computation~\cite{von2009human},
combine human and machine expertise with the machine as the ultimate
arbiter of behavior, and a group of humans acting as lower level
computational components.

True collaboration, however, does not start with one participant being
assigned to a leadership role.  Instead, the various agents need to
dynamically negotiate their roles and jockey for position, discovering
when and how to trust each other to move forward.  A concrete example
is in the context of self-driving cars.  Cars share the road with each
other and must carefully choose when to be responsive to other
vehicles and when to assert themselves to create a situation that
benefits them.  Doing so makes a significant difference in
effectiveness~\cite{cunningham2015mpdm}.  A related problem arises
when a self-driving wheelchair attempts to move through a group of
pedestrians~\cite{kim2016socially}.  More generally, robots that
interact with people in the physical world need to navigate the
complex give-and-take of establishing mutually beneficial behaviors
where possible.

Indeed, a wide variety of human--machine interaction problems would be
positively impacted by the technology we are proposing.
%
As such, the project fits well with Brown University's Humanity
Centered Robotics Initiative (HCRI), of which co-PI Littman is
co-director.  Specifically, within HCRI, there are ongoing efforts to
design robotic systems that interact with people and support
independent living tasks (e.g., gerontechnological support for aging
in place and robotic sous chefs) and our project will form an
important foundation for this broader work.
% For an elderly person (or chef) to trust and collaborate on tasks with
% a machine effectively, the machine must act in a manner that the
% elderly person (or chef) expects.  This research we are proposing to
% carry out is foundational for these important applications.

The most direct application of our ideas would be in scenarios in
which a robot is sharing a space with bystanders (human or robotic)
who do not necessarily share its goals---a personal grocery-store
assistant, a socially aware vacuum, an automated hospital linen cart, or
a package delivery drone.  To further develop our ideas in a physical
environment like these, and moreover in embodied robots,
% in year 3, team up with George and/or Stefanie. by then, there will
% be robots in the department that can both move and move things around
we plan to build a ``go fetch'' application on top of a
mobile-manipulation platform that is scheduled for deployment in the
Brown Computer Science department in Spring 2017.  In this
application, a user will ask a mobile maipulator robot in the
department to ``go fetch object X from person Y'' (for example,
``borrow the HCI textbook from Prof.\ Huang'' or ``pick up the package
that Genie has for me in the mail room'').  To be successful in this
capacity, the robot will need to be able to represent, learn, and
apply collaborative behavior in (at least) three separate spheres:
%
%\begin{itemize}
(1)~navigating smoothly through the atrium and hallways, finding
  appropriate ways to avoid collisions, and adopting regularities that
  make it easy for people to coordinate their own movements with it;
%
(2)~riding in the elevator, which brings with it some related but
  distinct expectations for coordinating (moving in and out of the
  doors, standing inside, asking for help with the buttons, deciding
  what to do when obstructed, etc.);
%
(3)~understanding how to best signal a handoff of the object it is
  fetching, interrupting people without being too timid or too
  boorish.
%\end{itemize}
%
%This task, is within our technological limitations (unlike wheelchairs or actual cars) 
%and would definitely result in some impact if our algorithms were able to help.

Interest in robotics and machine learning is growing within the Brown
CS department as elsewhere.  Presently, our faculty is designing a new
robotics introductory course sequence that includes classes on basic
robotics algorithms, hardware design, and human-centered evaluation of
robotics systems.  Inspired by the ideas in this proposal, we plan to
contribute a new intermediate robotics course for undergraduates
called ``Social autonomous driving.''  As a test platform, we will use
MIT's Duckietown platform,\footnote{\url{http://duckietown.mit.edu/}}
which has already been installed at Brown.  Students will develop
robots that drive around this test environment.  The main emphasis
will be on making sure the robotic drivers interact smoothly with
other robotic and remote-controlled cars.  The latter will be achieved
by endowing the robots with socially rational capabilities, thereby
enabling them to adapt to local ``customs'' 
%and driving styles
(like the ``Pittsburgh left'' and the ``Boston rotary'').

%%%

% In an effort to increase diversity in computer science, we have
% pre-selected two graduate students to participate in this
% project---one woman, and the other Hispanic.  All students (both
% graduate and undergraduate) who join our team will benefit from
% collaborating with the cognitive psychologists on our team.
% Specifically, they will strengthen their understanding of a number of
% fields, all of which are critical to the development of artificial
% agents that collaborate effectively with humans: e.g., behavioral
% economics, cognitive psychology, reinforcement learning, and software
% engineering.

The co-PIs have a strong record of mentoring students from
underrepresented groups, and intend to continue pursuing these efforts
in the context of this proposal.  In particular, we plan to integrate
our research on human--machine collaborations into Artemis, a free
summer program directed by co-PI Greenwald that introduces rising 9th
grade girls to computational thinking.  For example, we can have the
Artemis girls teach a robot to collaborate with them on routine tasks,
such as navigation or object search.

%folding laundry.
%\commenta{insert a different example!!!}
%\commenta{You know, girls should be doing housework.} 
% JLA: yeah, i was concerned about the folding laundry example for that reason... 

% Our deliverables include an open-source publicly accessible toolkit
% for agents to learn collaborative behavior via reinforcement
% learning. Further, we will build a database of machine-machine,
% human-machine, and human-human experimental results on
% collaborative-learning problems, which can serve as a benchmark for
% future researchers to build artifical agents that increasingly achieve
% human-like behavior.

Finally, we expect to publish the results of the proposed research in
top-tier archival, conference proceedings and journals with high
impact factors, and to present our work at innovative, non-archival
workshops (e.g., the AAAI symposia).

